Language learning via recursive pragmatic reasoning

### 2013-05-31 praglearn.tex
edit

\title{Language learning in context via recursive pragmatic reasoning}
% \title{Learning words via recursive pragmatic reasoning about other agents}
% : The blessing of social anxiety - not a huge fan of colon titles

@@

The goal of our current work is to investigate the possibilities for
integrating models of recursive pragmatic reasoning with models of
language learning, with the hope of capturing phenomena in both
domains. We begin by describing a proposal for bringing the two together, noting several issues in applying a purely normative framework to recursive reasoning under uncertainty. We next simulate findings
on pragmatic inference in one-shot games (replicating previous
work). We then build on these results to simulate the results of pragmatic
learning in the language acquisition setting where one communicator is
uncertain about the lexicon and in iterated communication games where both
communicators are uncertain about the lexicon.

@@

Experimental results suggest that communicators are able to establish
novel, consensus-based communication systems. For example, adults
playing a communication game using only novel symbols with no
conventional meaning will typically converge on a set of new
conventions which allow them to accomplish their task
\cite{galantucci2005}. Or in a less extreme example, communicators
asked to refer to novel objects invent conventional names for them
over the course of repeated interactions (e.g., ``the ice skater'' for
an abstract figure vaguely resembling an ice skater,
\cite{clark1986}). From a pure learning perspective this behavior is anomalous,
however: Since both agents know perfectly well that there is no
underlying convention to discover, there is nothing to learn from the
other's behavior. Furthermore, even if only one partner is producing
the novel expressions, their behavior still becomes more regular
(conventional) over time, which would seem to rule out a role for
learning---even if there is some pattern in the expressions the
speaker chooses to use, there is certainly nothing for the
\textit{speaker} to learn by observing these patterns, and thus their
behavior should not change over time.
% An obvious hack might be to simply stipulate that speakers take
% their own utterances as training data for further learning, but this
% would fail to explain yet another wrinkle in this phenomenon:
% speakers only regularize their production over time if they have
% some kind of feedback indicating that the listener understands; if
% such feedback is removed, the phenomenon is dramatically reduced (X X
% krauss 1966, branigan catchpole pickering 2011).

@@

A stronger example of how pragmatics can create communicative biases can be
observed by considering the Horn implicatures introduced in \S\ref{sec:horn-implicature}. 
In these games, pragmatic reasoning causes each agent to start out weakly biasedtowards the associations ``cheap'' $\leftrightarrow$
\textsc{common}, ``expensive'' $\leftrightarrow$ \textsc{rare}; this bias arises because their prior on literal meaning
is uniform and uninformative. A fully rational listener, therefore,
observing an uncertain speaker using this mapping, would discount it
as arising from this bias, and conclude that the speaker in fact did
not know which convention was in use. Conventional listeners,
however, believe that speakers do know which convention is in
use, and therefore tend to take their behavior as positive evidence that the
``good'' system is in use. Similarly, conventional speakers
will wager that since on average they will succed more often if
listeners are using the ``good'' system, they might as well try
it. When they succeed, they take their success as evidence that the listener was in
fact using the good system all along. As a result, dyads in this game
not only end up converging onto a stable system at a rate far above
chance (Fig.~\ref{fig:horn}.

@@

Our work identifies two biases in communicators' behavior: a tendency for pragmatic speakers
and listeners to accentuate useful, sparse patterns in their
communicative systems (\S\ref{sec:emergence}), and for short,
high-frequency (``low cost'') expressions to
be assigned to common objects (\S\ref{sec:horn-emergence}).
Both of these biases systematically drive the overall communicative system
towards greater global efficiency. In the long term, these processes
should leave their mark on the structure of the language itself, which
may contribute to explaining how languages become 
optimized for effective communication \cite{zipf1936,piantadosi2011}.

More generally, considering the interaction between pragmatics and learning may be a critical part of understanding human language. Our work here takes a first step towards unifying disparate strands of research that have treated language acquisition and language use as distinct.

>>>>

\title{Language learning via recursive pragmatic reasoning}
% \title{Learning words via recursive pragmatic reasoning about other agents}
% : The blessing of social anxiety - not a huge fan of colon titles

@@

The goal of our current work is to investigate the possibilities for
integrating models of recursive pragmatic reasoning with models of
language learning, with the hope of capturing phenomena in both
domains. We begin by describing previous models in each domain as well
as proposals for bringing the two together. We next simulate findings
on pragmatic inference in one-shot games (replicating previous
work). We then build on these results to simulate the results of pragmatic
learning in the language acquisition setting where one communicator is
uncertain about the lexicon and in iterated communication games where both
communicators are uncertain about the lexicon. We end by discussing the implications of our results for
work on language change.

@@

Experimental results suggest that communicators are able to establish
novel, consensus-based communication systems. For example, adults
playing a communication game using only novel symbols with no
conventional meaning will typically converge on a set of new
conventions which allow them to accomplish their task
\cite{galantucci2005}. Or in a less extreme example, communicators
asked to refer to novel objects invent conventional names for them
over the course of repeated interactions (e.g., ``the ice skater'' for
an abstract figure vaguely resembling an ice skater,
\cite{clark1986}). From a pure learning perspective this behavior is anomalous,
however: Since both agents know perfectly well that there is no
existing convention to discover, there is nothing to learn from the
other's behavior. Furthermore, even if only one partner is producing
the novel expressions, their behavior still becomes more regular
(conventional) over time, which would seem to rule out a role for
learning---even if there is some pattern in the expressions the
speaker chooses to use, there is certainly nothing for the
\textit{speaker} to learn by observing these patterns, and thus their
behavior should not change over time.
% An obvious hack might be to simply stipulate that speakers take
% their own utterances as training data for further learning, but this
% would fail to explain yet another wrinkle in this phenomenon:
% speakers only regularize their production over time if they have
% some kind of feedback indicating that the listener understands; if
% such feedback is removed, the phenomenon is dramatically reduced (X X
% krauss 1966, branigan catchpole pickering 2011).

@@

A stronger example of how pragmatics can create communicative biases can be
observed by considering the Horn implicatures introduced in \S\ref{sec:horn-implicature}. 
In these games, pragmatic reasoning causes each agent to start out weakly biasedtowards the associations ``cheap'' $\leftrightarrow$
\textsc{common}, ``expensive'' $\leftrightarrow$ \textsc{rare}; this bias arises because their prior on literal meaning
is uniform and uninformative. A fully rational listener, therefore,
observing an uncertain speaker using this mapping, would discount it
as arising from this bias, and conclude that the speaker in fact did
not know which convention was in use. Conventional listeners,
however, believe that speakers do know which convention is in
use, and therefore tend to take their behavior as positive evidence that the
``good'' system is in use. Similarly, conventional speakers
will wager that since on average they will succed more often if
listeners are using the ``good'' system, they might as well try
it. When they succeed, they take their success as evidence that the listener was in
fact using the good system all along. As a result, dyads in this game
 end up converging onto a stable system at a rate far above
chance (Fig.~\ref{fig:horn}).

@@

Our work identifies two biases in communicators' behavior: a tendency for pragmatic speakers
and listeners to accentuate useful, sparse patterns in their
communicative systems (\S\ref{sec:emergence}), and for short,
high-frequency (``low cost'') expressions to
be assigned to common objects (\S\ref{sec:horn-emergence}).
Remarkably, while both of these biases are locally sub-optimal for individual agents,
they both systematically drive the overall communicative system
towards greater global efficiency. In the long term, these processes
should leave their mark on the structure of the language itself, which
may contribute to explaining how languages become 
optimized for effective communication \cite{zipf1936,piantadosi2011}.
